### 2008-06-26 docs/paper/paper.tex
Post-Presentation commit
Various updates. WERTi will be restructured significantly this weekend.

\documentclass{article}

\usepackage[english]{babel}
\usepackage{qtree}
\usepackage{aleks}
\usepackage[utf8x]{inputenc}
\usepackage{linguex}

\author{Aleksandar Dimitrov}
\title{WERTi - Easing Second Language Acquisition}

\begin{document}
\maketitle

\tableofcontents

\section*{Preface}

\section{Introduction}

Using tools and methods made available through new achievements in computational
linguistics and related subjects like cognitive science to ease the process of
second language acquisition has only recently gained focus in research projects.
While adaptions of traditional data sources like dictionaries have been used for
quite a while we have yet to discover new and effective ways of

WERTi tries to approach this probmlem from a more general perspective. Making
use of the the momentum of the Internet, WERTi provides a platform for
implementing linguistic analysis and subsequent input enhancement methods on
user specified pages on the World Wide Web. Using Java Servlets technology for
serving content and the UIMA framework for processing it in a dynamic and
flexible session, we aim to provide a platform for linguistic processing of
online content that can go beyond input enhancement.

\subsection{Acknowledgements}

I'd like to thank my advisor \textsc{Detmar Meurers} for providing me with the
unique opportunity on working on the WERTi system. WERTi was originally thought
out and developed in Python by him and his team at the Ohio State University. My
job was to port the system he had already written to Java using UIMA and
generalize the concept as to make it easier to implement new functions.

I would also like to thank \textsc{Janina Radó}, who provided me with invaluable feedback
about the general style guidelines of papers of this kind.

Furthermore I'd like to thank , , , and for providing testing the system and for
their suggestions, remarks and feature requests.

\section{Explaining the Core Functionality}

This section will explain the principles underlying WERTi's functionality by
first looking at the data processing architecture and then showing how this is
then integrated into the user interface of the web application.

\subsection{UIMA Analysis Engines}

The current processing line can be split up into three different parts that can
interact autonomously but each depend on the output of their predecessor. The
components are arranged in the following order:

\begin{itemize}

\item[HTML processing:] annotates HTML tags, finds ``relevant'' text within the
tags. This lays ground to later linguistic analysis by setting the margins of
where the latter has to operate on in the web content.

\item[Linguistic processing:] performs linguistic tasks (tokenization, sentence
boundary detection, part-of-speech tagging\ldots{}) on the text-annotations
from the previous processing step. This is also the most intensive step from a
computational point of view. Optimizations to the code are most likely to yield
visible results here.

\item[Post processing:] Performs all sorts of content enhancement on the
annotations of the previous two processing steps. Note that this step depends on
both annotation results: it also needs certain HTML markups in order to
correctly organize all of the code that is going to be executed on client side.

\end{itemize}

We will now go on and explain those steps in further detail.

\subsubsection{HTML Processing}

The HTML processor method was designed to be primitive and efficient. While
using a fully capable HTML parser was considered, we preferred to use a more
simple approach. Full and formally correct HTML markup was deemed unnecessary and
too time intensive.
Furthermore, changing the
implementation of an analysis step even this fundamental to further processing
should be easy and without side-effects as long as the requirements to
preconditions and postconditions are met.

\paragraph{Preconditions} We have a document represented as a singleton
String\footnote{As usual in UIMA. For large documents, UMIA provides ways of
splitting up documents and processing the chunks independently. However, UIMA
only considers documents well beyond one megabyte to be ``big enough'' to be
split and the plain HTML most web pages serve rarely exceeds this mark.}   and
stored in the UIMA-CAS that is passed to the \verb#process(CAS)#-method of the
\verb#HTMLAnnotator#-class.
\paragraph{Postconditions} The document contains annotations marking up the
positions and spans of html tags in the document text. The names of
tags\footnote{E.g. ``p'' for the tag \texttt{<p>} or ``div'' for the dag
\texttt{div}.} are
also stored and a \verb#isClosing#-flag is set, that denotes whether this tag is
closing another\footnote{A preceding slash in the tag name closes the tag.}.

\subsubsection{Linguistic Processing}

Linguistic processing goes through 3 major steps: \emph{Tokenization},
\emph{Sentence Boundary Detection}

\subsubsection{Post Processing - Input Enhancement}

\section{The User Interface}

\subsection{Web Interface}

\subsection{Programming Interface}

\section{The Development Process}
In this section I want to display in what environment the project has been
written and what technologies have been put to use.

\subsection{Concept \& Design}
The original design of WERTi has been developed at Ohio State University by
\textsc{Detmar Meurers} and his research associates. There exist several papers
about this original work. At Tübingen University the concept has been extended
to

\subsection{Design Process} My Work on the Java implementation of WERTi was
mostly autonomous. I had been given free choice over the frameworks and
development environment I wanted to put to use, with the only exceptions being
being using \emph{Java} and the UIMA architecture to achieve maximum flexibility
and portability. Maximum flexibility and extensibility were the main goals of
the re-implementation.

Development happened in a very productive atmosphere with weekly project
meetings where core functionality and the design of the analysis process were
discussed. After those most important aspects of the system were implemented, I
continued work on the interface and extension of the analysis process mostly
alone for organizational reasons.

I chose to use mostly standard UNIX command line tools for writing, testing and
debugging the code, although more advanced solutions like the Eclipse Java IDE
were also used - mostly for easier organization of the work with the different
frameworks\footnote{Especially UIMA which comes with a number

\section{Conclusion}

\subsection{Loose Ends}

\end{document}